The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-07-31 WavesData-2019/LatexNotes/notes.tex
NM

\begin{comment}
{\bf Sidenote}: In the case of constant depth, if one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
\begin{equation}
2 u_t + 3 \eps u u_x + \frac{\delta^2}{3} u_{xxx} = 0
\end{equation}
where $\eps = a/h$ and $\delta = h/\lam$.
\end{comment}

>>>>

%\begin{comment}
{\bf Sidenote}: In the case of constant depth, if one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
\begin{equation}
2 u_t + 3 \eps u u_x + \frac{\delta^2}{3} u_{xxx} = 0
\end{equation}
where $\eps = a/h$ and $\delta = h/\lam$.
%\end{comment}

### 2019-07-31 paper2_JNLS/paper.tex
NM

Here, we recast the variable-depth KdV equation \eqref{KdV} in dimensionless variables.
Of course, the choice of how to nondimensionalize is not unique, and our choices will made for convenience with the Hamiltonian structure introduced in the next section. It may be instructive to be able to compare against other traditional nondimensionalizations and hence we first introduce a generic nondimensionalization: nondimensionalize 
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \xi / \lengthscale
&&\mbox{\em dimensionless position (moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
where the characteristic scales $\ampscale, \lengthscale, \timescale$ for wave amplitude, length, and time respectively, are free for us to choose.
Then the dimensionless KdV equation is
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}

@@

In particular, we make the following choices
\begin{align}
&\ampscale = \pi^{1/2} \, \etastd \\
&\lengthscale = \lamfac \lam / \pi \\
&\timescale = \lamfac / \freqp
\end{align}
where $\lamfac$ is an integer to be chosen later. In the physical domain, we will consider an integer multiple of the characteristic wavelength $\xi \in [-\lamfac \lam, \lamfac \lam]$, so that the state variable $u$ is periodic over the  dimensionless domain $\tilde{x} \in [-\pi, \pi]$. We have scaled time accordingly. Finally, our choice of the scale for wave amplitude $\ampscale$ is made to normalize the energy of $u$ to unity. The Hamiltonian framework that we will use relies heavily on the energy, and hence it is more convenient to have the energy normalized to unity rather than making the more common choice of $\etastd$ by itself as the scale for wave amplitude.

With these choices, the dimensionless KdV is
\begin{align}
\label{dimlessKdV}
&u_t + C_3 \drat^{-1} \, u u_x + C_2 \drat \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi] \\
&C_3 = \frac{3}{2} \pi^{3/2} \epsup  \\
&C_2 = \frac{\pi^3 \delup^2}{6 \lamfac^2} 
\end{align}
where we have the dimensionless parameters
\begin{align}
&\epsup = \etastd / \dup
&&\mbox{\em upstream amplitude-to-depth ratio} \\
&\delup = \dup / \lamup
&&\mbox{\em upstream depth-to-wavelength ratio} \\
&\drat = {\depth}/{\dup} \, .
&&\mbox{\em dimensionless depth}
\end{align}
Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC. In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$. Then we have
\begin{equation}
\drat = 
\begin{cases}
1 		&\quad \mbox{for } t<0 \\
\dratdn 	&\quad \mbox{for } t>0
\end{cases}
\end{equation}
where $\dratdn = \ddn/\dup$.
See Table \ref{paramtable} for a summary of important parameters.

>>>>

Here, we recast the variable-depth KdV equation \eqref{KdV} in dimensionless variables.
Of course, the choice of how to nondimensionalize is not unique, and our choices will made for convenience with the Hamiltonian structure introduced in the next section. It may be instructive to to compare against other traditional nondimensionalizations and hence we first introduce a generic nondimensionalization:
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \xi / \lengthscale
&&\mbox{\em dimensionless position (moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
where the characteristic scales $\ampscale, \lengthscale, \timescale$ for wave amplitude, length, and time respectively, are free for us to choose.
Then the dimensionless KdV equation is
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}

@@

In particular, we make the following choices
\begin{align}
\ampscale = \pi^{1/2} \, \etastd \, , \qquad
\lengthscale = \frac{\lamfac \lam}{\pi} \, , \qquad
\timescale = \frac{\lamfac \lam}{\freqp \depth}
\end{align}
where $\lamfac$ is an integer to be chosen later. In the physical domain, we will consider an integer multiple of the characteristic wavelength $\xi \in [-\lamfac \lam, \lamfac \lam]$, so that the state variable $u$ is periodic over the  dimensionless domain $\tilde{x} \in [-\pi, \pi]$. We have scaled time accordingly. Finally, our choice of the scale for wave amplitude $\ampscale$ is made to normalize the energy of $u$ to unity. The Hamiltonian framework that we will use relies heavily on the energy, and hence it is more convenient to have the energy normalized to unity rather than making the more common choice of $\etastd$ by itself as the scale for wave amplitude.

With these choices, the dimensionless KdV is
\begin{align}
\label{dimlessKdV}
&u_t + C_3 \drat^{-3/2} \, u u_x + C_2 \drat^{1/2} \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi] \\
&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1}  \\
&C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
\end{align}
where we have the dimensionless parameters
\begin{align}
&\epsup = \etastd / \dup
&&\mbox{\em upstream amplitude-to-depth ratio} \\
&\delup = \dup / \lamup
&&\mbox{\em upstream depth-to-wavelength ratio} \\
&\drat = {\depth}/{\dup} \, .
&&\mbox{\em dimensionless depth}
\end{align}
Here, $\drat = \dratupdn$ is the local dimensionless depth, which changes value crossing the ADC. In particular, since \eqref{dimlessKdV} is in the reference frame that moves with the characteristic wave speed, we can assume that the ADC is met at $t = T_{ADC} = 0$. Then we have
\begin{equation}
\drat = 
\begin{cases}
1 		&\quad \mbox{for } t<0 \\
\dratdn 	&\quad \mbox{for } t>0
\end{cases}
\end{equation}
where $\dratdn = \ddn/\dup$.
See Table \ref{paramtable} for a summary of important parameters.
